---
title: "Assignment 1"
format: html
editor: visual
author: Namra Shahid
author2: Marit Rygg Fredheim
bibliography: 
csl: apa-no-ampersand.csl
---

## Describe the data (GDP per capita)

```{r}
library(tidyverse)
library(vtable)
library(dineq)
library(dplyr)
```

```{r}
library(readr)
gdpfile <- read.csv('nama_10r_3gdp.csv')
populationfile <- read_csv("demo_r_pjanaggr3.csv")

gdpdata <- gdpfile %>%
  rename(Year = TIME_PERIOD, GDP = OBS_VALUE, Region = geo)

populationdata <- populationfile %>% 
  rename(Year = TIME_PERIOD, Population = OBS_VALUE, Region = geo)
```

## GDP per capita calculation:

We are here calcualting GDP per capita:

Statistically explained:

**yi= GDPi/ populationi**

```{r}
# Calculate GDP per capita
tmp <- gdpdata %>%
  #combine gdpr data with population data with 
  # the same year and country
  left_join(populationdata, by = c("Region", "Year")) %>% 
  select(Region, Year, GDP, Population) %>% 
  mutate(
    GDP_capita = (GDP * 1000000)/Population
  )
```

## Report the essential descriptive statistic

```{r}
# Report the essential descriptive statistics 
# for you data set (Mean, Min, Max,..:)
tmp %>% 
  select(GDP, Population) %>% 
  # . to stop RStudio complaining about no data
  vtable::st(.)
```

```{r}
#NUTS2 files
library(readr)
gdpfile2 <- read.csv('nama_10r_3gdp_nuts2.csv')
populationfile2 <- read_csv("demo_r_pjanaggr3_nuts2.csv")

gdpdata2 <- gdpfile2 %>%
  rename(Year = TIME_PERIOD, GDP = OBS_VALUE, Region = geo)

populationdata2 <- populationfile2 %>% 
  rename(Year = TIME_PERIOD, Population = OBS_VALUE, Region = geo)
```

```{r}
# Calculate GDP per capita
tmp_ginie <- gdpdata2 %>%
  #combine gdpr data with population data with 
  # the same year and country
  left_join(populationdata2, by = c("Region", "Year")) %>% 
  select(Region, Year, GDP, Population) %>% 
  mutate(
    GDP_capita2 = (GDP * 1000000)/Population
  )
```

```{r}
# Skal nok vekk


# Report the essential descriptive statistics 
# for you data set (Mean, Min, Max,..:)
tmp22 %>% 
  select(GDP, Population) %>% 
  # . to stop RStudio complaining about no data
  vtable::st(.)
```

```{r}
kable(summary(tmp_ginie[,3:5]))
```

```{r}
tmp_ginie$GDP_capita <- as.numeric(tmp_ginie$GDP_capita)
tmp_ginie$Population <- as.numeric(tmp_ginie$Population)
```

```{r}
gini.wtd(tmp_ginie$GDP_capita, weights = tmp_ginie$Population)
```

```{r}
tmp_ginie <- tmp_ginie %>%
  mutate(NUTS2 = substr(tmp_ginie$Region,1,4))

tmp_ginie <- tmp_ginie %>%
  mutate(NUTS = substr(tmp_ginie$Region,1,2))

na.omit(tmp_ginie)

tmp_ginie %>% 
  group_by(NUTS2) %>%
  mutate(GINI_NUTS2 = gini.wtd(GDP_capita, weigths = Population)) %>%
  ungroup()
```

Nuts 2

## Discuss briefly if there are noteworthy outliers:

## 

An important thing to come up here is that an outlier is a value or an observation that is distant from another observations, that is to say, a data point that differs significantly from other observations.

An observation must always be compared to other observations made on the same phenomenon.
In this assignment the focus is gdp per capita and gdp.
