---
title: "Assignment 1 (Data and Descriptive)"
format: html
editor: visual
author: Namra Shahid
author2: Marit Rygg Fredheim
bibliography: 
csl: apa-no-ampersand.csl
---

In this assignment we are going to Calculate GDP per capita for 8 subset of countries sub-regional (NUTS3).
We are group two and will therefore examine these countries:

-   Ireland (IE)

-   Poland (PL)

-   Spain (ES)

-   Slovania (SI)

-   Netherlands (NL)

-   Greece (EL)

-   Norway (NO)

-   Albania (AL

## Sub-national GDP

```{r}
library(tidyverse)
library(vtable)
library(dineq)
library(dplyr)
```

```{r}
library(readr)
gdpfile <- read.csv('nama_10r_3gdp.csv')
populationfile <- read_csv("demo_r_pjanaggr3.csv")

gdpdata <- gdpfile %>%
  rename(Year = TIME_PERIOD, GDP = OBS_VALUE, Region = geo)

populationdata <- populationfile %>% 
  rename(Year = TIME_PERIOD, Population = OBS_VALUE, Region = geo)
```

### GDP per capita calculation:

We are here calcualting GDP per capita:

Statistically explained:

**y*i*= GDP*i*/ population*i***

```{r}
# Calculate GDP per capita
tmp <- gdpdata %>%
  #combine gdpr data with population data with 
  # the same year and country
  left_join(populationdata, by = c("Region", "Year")) %>% 
  select(Region, Year, GDP, Population) %>% 
  mutate(
    GDP_capita = (GDP * 1000000)/Population
  )
```

```{r}
#Legg inn ein tabell for gdp per capita
ggplot(data = tmp)
```

**Description of the data!**

### Report the essential descriptive statistic

```{r}
# Report the essential descriptive statistics 
# for you data set (Mean, Min, Max,..:)
tmp %>% 
  select(GDP, Population) %>% 
  # . to stop RStudio complaining about no data
  vtable::st(.)
```

## Regional inequity

The first files only contains data on NUTS3 level.
In this part of the assignment we need data on NUTS2 level, therefore we need a new file.
For the next submission, we can download a new data file with all data available, if needed.

```{r}
#NUTS2 files
library(readr)
gdpfile2 <- read.csv('nama_10r_3gdp_nuts2.csv')
populationfile2 <- read_csv("demo_r_pjanaggr3_nuts2.csv")

gdpdata2 <- gdpfile2 %>%
  rename(Year = TIME_PERIOD, GDP = OBS_VALUE, Region = geo)

populationdata2 <- populationfile2 %>% 
  rename(Year = TIME_PERIOD, Population = OBS_VALUE, Region = geo)
```

```{r}
# Calculate GDP per capita for NUTS2
tmp_ginie <- gdpdata2 %>%
  #combine gdpr data with population data with 
  # the same year and country
  left_join(populationdata2, by = c("Region", "Year")) %>% 
  select(Region, Year, GDP, Population) %>% 
  mutate(
    GDP_capita2 = (GDP * 1000000)/Population
  )
```

### Essential descriptive statistics for NUTS2

```{r}
# Report the essential descriptive statistics 
# for you data set (Mean, Min, Max,..:)
tmp22 %>% 
  select(GDP, Population) %>% 
  # . to stop RStudio complaining about no data
  vtable::st(.)
```

```{r}
kable(summary(tmp_ginie[,3:5]))
```

```{r}
tmp_ginie$GDP_capita <- as.numeric(tmp_ginie$GDP_capita)
tmp_ginie$Population <- as.numeric(tmp_ginie$Population)
```

```{r}
gini.wtd(tmp_ginie$GDP_capita, weights = tmp_ginie$Population)
```

```{r}
tmp_ginie <- tmp_ginie %>%
  mutate(NUTS2 = substr(tmp_ginie$Region,1,4))

tmp_ginie <- tmp_ginie %>%
  mutate(NUTS = substr(tmp_ginie$Region,1,2))

na.omit(tmp_ginie)

tmp_ginie %>% 
  group_by(NUTS2) %>%
  mutate(GINI_NUTS2 = gini.wtd(GDP_capita, weigths = Population)) %>%
  ungroup()
```

```{r}
kable(summary(tmp_ginie[8]))
```

```{r}
ggplot(data = tmp_ginie) + geom_density(aes(x = GINI_NUTS2), 
fill = "maroon", col = "maroon", alpha = 0,2) + theme_bw()
```

## Discuss briefly if there are noteworthy outliers:

An important thing to come up here is that an outlier is a value or an observation that is distant from another observations, that is to say, a data point that differs significantly from other observations.

An observation must always be compared to other observations made on the same phenomenon.
In this assignment the focus is gdp per capita and gdp.
